---
title: "pipe"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
    css: config/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
```

## Why pipe

You can use the pipe to rewrite multiple operations that you can read left-to-right, top-to-bottom (reading the pipe operator as "then").

* too many variables:

    ```{r eval=FALSE}
    n1 <- rnorm(10000)
    n2 <- abs(n1)
    n3 <- matrix(n2, ncol = 100)
    n4 <- rowMeans(n3)
    n5 <- round(n4)
    hist(n5)
    ```

* nested functions:

    ```{r eval=FALSE}
    hist(round(rowMeans(matrix(abs(rnorm(10000)), ncol = 100))))
    ```

* more readable with `%>%`!

    ```{r eval=FALSE}
    rnorm(10000) %>% 
      abs %>% 
      matrix(ncol = 50) %>% 
      rowMeans() %>% 
      round %>% 
      hist
    ```

Pipe make your code more readable by:

* structuring sequences of data operations left-to-right (as opposed to from the inside and out),
* avoiding nested function calls,
* minimizing the need for local variables and function definitions, and
* making it easy to add steps anywhere in the sequence of operations.

## Usage

### Basic piping
  
* `x %>% f` is equivalent to `f(x)`
* `x %>% f(y)` is equivalent to `f(x, y)`
* `x %>% f %>% g %>% h` is equivalent to `h(g(f(x)))`

### The argument placeholder

* `x %>% f(y, .)` is equivalent to `f(y, x)`
* `x %>% f(y, z = .)` is equivalent to `f(y, z = x)`
 
### Re-using the placeholder for attributes

* `x %>% f(y = nrow(.), z = ncol(.))` is equivalent to 
   `f(x, y = nrow(x), z = ncol(x))`

* `x %>% {f(y = nrow(.), z = ncol(.))}` is equivalent to 
   `f(y = nrow(x), z = ncol(x))`

## Examples

```{r}
instancename <- c('Cycle 1 Day 1', 'Cycle 2 Day 1')

folder <- instancename %>% 
  str_replace_all('Cycle (\\d+) Day (\\d+)', 'C\\1D\\2')

folder
```

```{r}
starwars %>% 
  select(name, height, species) %>% 
  filter(species %in% c('Human', 'Droid')) %>% 
  mutate(mean_height = mean(height, na.rm = T)) %>% 
  group_by(species) %>% 
  mutate(mean_height_by_species = mean(height, na.rm = T)) %>% 
  ungroup()
```

## Reference

[GitHub - tidyverse/magrittr](https://github.com/tidyverse/magrittr){target="_blank"}

[R语言高效的管道操作magrittr](http://blog.fens.me/r-magrittr/){target="_blank"}

