---
title: "stringr"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
    css: [config/style.css, config/multi-columns.css]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = T,
                      eval = TRUE,          # run (or not) the code in the chunk
                      echo = TRUE,          # display (or not) the code of the chunk
                      message = TRUE,       # display (or not) the messages of the chunk
                      warning = TRUE,       # display (or not) the warning message of the chunk
                      error = FALSE,        # display (or not) the error message of the chunk
                      collapse = TRUE,      # collapse (or not) all the source and output blocks in to a single block
                      comment = '#>'        # the start of each line of results in the final document)
)
```

```{r message=FALSE}
library(tidyverse)
```

## Overview

__stringr__ is built on top of __stringi__, which uses the ICU C library to provide fast, correct implementations of common string manipulations. stringr focusses on the most important and commonly used string manipulation functions whereas stringi provides a comprehensive set covering almost anything you can imagine. If you find that stringr is missing a function that you need, try looking in stringi. Both packages share similar conventions, so once you’ve mastered stringr, you should find stringi similarly easy to use.

#### generate data
```{r}
fruit <- stringr::fruit[1:30]

sentences <- stringr::sentences[1:20]

foo <- c('   a   bc   d   ')
```

## Regular Expressions

#### `str_view(string, pattern, match = NA)`
#### `str_view_all(string, pattern, match = NA)`

```{r, collapse=FALSE}
str_view(fruit[1:5], '[aeiou]')
str_view_all(fruit[1:5], '[aeiou]')
```

#### special character---------`[ ]` `\` `^` `$` `.` `|` `?` `*` `+` `( )`

* `\` (learn more about escaped characters: `?'"'`)
    ```{r, collapse=FALSE}
    quote <- c(" ' ", ' " ', ' \' ', " \" ")
    writeLines(quote)
    
    str_view_all(c("ab\nc", "12d", "ae2"), "\n")
    str_view_all(c("ab\nc", "12\bd", "ae2"), "\b")
    str_view_all(c("ab\nc", "12\bd", "ae2"), "\\b")
    str_view_all(c("ab\nc", "12\\d", "ae2"), "\\\\")
    ```

* `\d` matches any digit(`\D`)
* `\s` matches any whitespace (e.g. space, tab, newline)(`\S`)
* `\w` match any word character,which includes alphabetic characters, marks and decimal numbers(`\W`)
* `\b` matches word boundaries, the transition between word and non-word characters(`\B`)

    ```{r, collapse=FALSE}
    str_view_all(c("abc", "12d", "ae2"), "\\d")
    str_view_all(c("abc", "\\12d", "1e2"), "\\w")
    str_view_all(c("a c", "\\1\td", "1e2"), "\\s")
    str_view_all(c("a c", "\\1\td", "1e2"), "\\b")
    ```


* `| [] ()`
    ```{r, collapse=FALSE}
    str_view(c("longest", "lonlest", "lonaest"), "lon(g|l)est")
    str_view(c("longest", "lonlest", "lonaest"), "lon[gal]est")
    str_view(c("longest", "lonlest", "lonaest"), "lon[a-g]est")
    str_view(c("longest", "lonlest", "lonaest"), "lon[^a-g]est")
    ```

* `.` matches any character (except a newline)
    ```{r, collapse=FALSE}
    x <- c("apple", "ba\nnana", "pear")
    writeLines(x)
    str_view_all(x, ".a.")
    str_view_all(x, ".a\\b")
    str_view_all(x, ".a\\n")
    ```

* `^`/`$` match the start/end of the string
    ```{r, collapse=FALSE}
    str_view(c("a\\abpl^e", "bbna\bna", "pear", "aaa"), "^a")
    str_view(c("apple", "banana", "pear"), "a$")
    str_view(c("$^$"), "\\$\\^\\$")
    ```

* Repetition

  `?`: 0 or 1
  
  `+`: 1 or more
  
  `*`: 0 or more
  
  `{n}`/`{n,}`/`{n,m}`
  
    ```{r, collapse=FALSE}
    str_view("1888 is the longest year in Roman numerals: MDCCCLXXXVIII", "CC?")
    str_view("1888 is the longest year in Roman numerals: MDCCCCLXXXVIII", "CC+")
    str_view("1888 is the longest year in Roman numerals: MDCCCLXXXVIII", "CC*")
    str_view("1888 is the longest year in Roman numerals: MDCCCLXXXVIII", "C{2,3}")
    ```

* Grouping and backreferences
    ```{r, collapse=FALSE}
    str_view(fruit[16:25], "(..)\\1")
    str_view(c("bacdb","bacdbacd"), "^(.).*\\1$")
    ```

#### `regex(pattern, ignore_case = FALSE, multiline = FALSE, comments = FALSE, dotall = FALSE)`

* `ignore_case`:	Should case differences be ignored in the match?
* `multiline`:  If `TRUE`, "$" and "^" match the beginning and end of each line.
                If `FALSE`, the default, only match the start and end of the input.
* `comments`:  If `TRUE`, white space and comments beginning with "#" are ignored. Escape literal spaces with "\\".
* `dotall`:  If `TRUE`, "." will also match line terminators.

```{r, collapse=FALSE}
x <- c("apple", "ba\nnana", "pear")

str_view_all(x, regex(".A.", ignore_case=T, dotall = T))
str_view_all("A\nb", regex("^b", multiline=T))
```


## Detect Matches
* `str_detect(string, pattern, negate = FALSE)`:	Detect the presence or absence of a pattern in a string.
* `str_which(string, pattern, negate = FALSE)`: find positions.
* `str_count(string, pattern = "")`: Count the number of matches in a string.
* `str_locate(string, pattern)`: returns an integer matrix
* `str_locate_all(string, pattern)`: returns a list of integer matrices

```{r}
fruit
```

#### `str_detect(string, pattern, negate = FALSE)`
```{r}
str_detect(fruit[1:5], 'a', negate = T)

test <- matrix(fruit[1:10],nrow=5)
str_detect(test,'a')

str_detect('appleapricot', fruit[1:5])

str_detect(fruit[1:5], c('a', 'b', 'c', 'd', 'e'))
```

#### `str_which(string, pattern, negate = FALSE)`
```{r}
str_which(fruit[1:5], 'a')

str_which(fruit, 'a', negate = T)
```

#### `str_count(string, pattern = "")`
```{r}
str_count(fruit[1:5], 'a')

str_count('abababa', 'aba')
```
#### `str_locate(string, pattern)`

#### `str_locate_all(string, pattern)`
```{r}
str_locate(fruit[1:5], 'a')

str_locate_all(fruit[1:5], 'a')
```

## Subset Strings
* `str_sub(string, start = 1L, end = -1L)`  Extract and replace substrings from a character vector.
* `str_subset(string, pattern, negate = FALSE)` Keep strings matching a pattern, or find positions.
* `str_extract(string, pattern)` Extract matched groups from a string.
* `str_extract_all(string, pattern, simplify = FALSE)` Extract matched groups from a string.
```{r}
fruit
```

#### `str_sub(string, start = 1L, end = -1L)`
```{r}
str_sub(fruit[1:5], 1, 3)

str_sub(fruit[1:5], end = -2)

str_sub(fruit[1:5], -2)

str_sub(fruit[1:5], c(1, 2, 3, 1, 1), c(3, 3, 4, 3, 3))

# str_locate(fruit[1:5], '(..)\\1')
# str_sub(fruit[1:5], str_locate(fruit[1:5], '(..)\\1'))
```

#### `str_subset(string, pattern, negate = FALSE)`
```{r}
str_subset(fruit, 'b')

# fruit[str_detect(fruit,'b')]
```

#### `str_extract(string, pattern)`
#### `str_extract_all(string, pattern, simplify = FALSE)`
```{r}
str_extract(c('aeaebcbc','bcaebc','abcccc'),'(..)\\1')
str_extract(sentences, '\\w+(es|s)\\b')

str_extract_all(c('aeaebcbc','bcaebc','abcccc'),'(..)\\1')

str_extract_all(c('aeaebcbc','bcaebc','abcccc'),'(..)\\1', simplify = T)

```

#### `str_match(string, pattern)` Extract matched groups from a string.
#### `str_match_all(string, pattern)`
```{r}
sentences[1:10]

str_match(sentences[1:10], '(a|the) ([^ ]+)')

# str_extract(sentences[1:10], '(a|the) ([^ ]+)')
# str_match(sentences[1:10], '(a|the) ([^ ]+)')[,1]

str_match_all(sentences[1:5], '(a|the) ([^ ]+)')
```

## Mutate Strings
* `str_sub(string, start = 1L, end = -1L, omit_na=FALSE) <- value`
* `str_replace(string, pattern, replacement)`
* `str_replace_all(string, pattern, replacement)`
* `str_remove(string, pattern)`
* `str_remove_all(string, pattern)`
* `str_to_upper(string, locale = "en")`
* `str_to_lower(string, locale = "en")`
* `str_to_title(string, locale = "en")`
```{r}
fruit
fruit_temp <- fruit
```

#### `str_sub(string, start = 1L, end = -1L, omit_na=FALSE) <- value`
```{r}
str_sub(fruit_temp,1,3) <- 'str'
fruit_temp
```

#### `str_replace(string, pattern, replacement)`
#### `str_replace_all(string, pattern, replacement)`
```{r}
str_replace(fruit[1:5], 'a', '-')

str_replace_all(fruit[1:5], 'a', '-')
```

#### `str_remove(string, pattern)`
#### `str_remove_all(string, pattern)`
```{r}
example <- c("apple", "apap", "ap")
str_remove(example, 'ap') 

str_remove_all(example, 'ap')

str_remove_all(example, '[ap]')
```

#### `str_to_upper(string, locale = "en")`
#### `str_to_lower(string, locale = "en")`
#### `str_to_title(string, locale = "en")`
```{r}
sentences[1:5]

str_to_upper(sentences[1:5])

str_to_lower(sentences[1:5])

str_to_title(sentences[1:5])
```

## Join and Split

join

* `str_c(..., sep = "", collapse = NULL)`

* `str_dup(string, times)`

* `str_glue(...)`

* `str_glue_data(.x, ..., .na = "NA")`

Split

* `str_split(string, pattern, n = Inf, simplify = FALSE)`

* `str_split_fixed(string, pattern, n)`

`str_c(..., sep = "", collapse = NULL)`:

```{r}
letters[1:5]

LETTERS[1:5]

# for one part
str_c(letters[1:5], collapse =' ')

# for two or more part
str_c(letters[1:5], LETTERS[1:5], sep = '/')

str_c(letters[1:5], LETTERS[1:5], sep = '/', collapse = ' ')

```

`str_dup(string, times)`:

```{r}
str_dup(fruit[1:6], times = 2)

str_dup(fruit[1:6], 1:3)
```

#### `str_glue(...)`

#### `str_glue_data(.x, ..., .na = "NA")`
```{r}
name <- 'Bob'
age <- 50
anniversary <- as.Date("1991-10-12")
str_glue(
  "My name is {name}, ",
  "my age next year is {age + 1}, ",
  "and my anniversary is {format(anniversary, '%A, %B %d, %Y')}."
)

str_glue(
  "My name is {name}, ",
  "and my age next year is {age + 1}.",
  name = "Joe",
  age = 40
)

str_glue('the {i}th letter is {letters[1:5]}', i = 1:5)

# here str_glue() is more readable than str_c()
str_c('the ', 1:5, 'th letter is ', letters[1:5])

str_glue('the {i}th {{letter}} is {{{letters[1:5]}}', i=1:5)

str_glue_data(mtcars[1:5, ], '{hp} hp')

mtcars[1:5, ] %>% str_glue_data('the {rownames(.)} has {hp} hp')
```

#### `str_split(string, pattern, n = Inf, simplify = FALSE)`
#### `str_split_fixed(string, pattern, n)`

`str_split(string, pattern, simplify = T)` is equivalent to `str_split_fixed(string, pattern, n = Inf)`

```{r}
fruits_split <- c(
  "apples and oranges and pears and bananas",
  "pineapples and mangos and guavas"
)

str_split(fruits_split, " and ")

str_split(fruits_split, " and ", simplify = TRUE)


# Specify n to restrict the number of possible matches
str_split(fruits_split, " and ", n = 3)

str_split(fruits_split, " and ", n = 2)

# If n greater than number of pieces, no padding occurs
str_split(fruits_split, " and ", n = 5)


# Use str_split_fixed() to return a character matrix
str_split_fixed(fruits_split, " and ", 3)

str_split_fixed(fruits_split, " and ", 4)

```

## Manage Lengths

* `str_length(string)`
* `str_pad(string, width, side = c("left", "right", "both"), pad =" ")` Pad a string
* `str_trunc(string, width, side = c("right", "left", "center"), ellipsis = "...")` Truncate a character string.
* `str_trim(string, side = c("both", "left", "right"))` Trim whitespace from a string
* `str_squish(string)`

```{r}
fruit
```

#### `str_length(string)`
```{r}
x <- c('adv', 'ss', 'awsd')
writeLines(x)

length(x)

str_length(x)
```

#### `str_pad(string, width, side = c("left", "right", "both"), pad =" ")` Pad a string
```{r}
str_pad(c("a", "abc", "abcdef"), 10, side = 'right')

str_pad("a", c(5, 7, 10))

str_pad("a", 10, pad = c("-", "_", " "))

```

#### `str_trunc(string, width, side = c("right", "left", "center"), ellipsis = "...")` Truncate a character string.
```{r}
str_trunc(fruit, 7)

str_trunc(fruit, 7, side = 'left', ellipsis = '**')
```

#### `str_trim(string, side = c("both", "left", "right"))` Trim whitespace from a string
#### `str_squish(string)`
```{r}
str_trim(c('    a', 'b     ', 'c   d'))

str_trim(c('    a', 'b     ', 'c   d'), side = 'left')

str_squish(c('    a', 'b     ', 'c   d'))

#str_replace_all(c('    a', 'b     ', 'c   d'), '\\s', '')
```

## Order Strings

* `str_order(x, decreasing = FALSE, na_last = TRUE, locale = "en", numeric = FALSE, ...)`
* `str_sort(x, decreasing = FALSE, na_last = TRUE, locale = "en", numeric = FALSE, ...)`

```{r}
bar <- c('b', 'c', 'a', 'd')

str_order(bar)

bar[str_order(bar)]

x <- c("100a10", "100a5", "2b", "2a")
str_sort(x)
str_sort(x, numeric = TRUE)

```


## Exercise
1. Use `str_length()` and `str_sub()` to extract the middle character from a string. What will you do if the string has an even number of characters?
2. Find all words that start with a __vowel__ and end with a __consonant__ (words).


## Helpers
More details about regular expressions:
[CRAN | stringr/Regular expressions](https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html){target="_blank"}


## Reference

[cheatsheet - stringr](https://github.com/rstudio/cheatsheets/blob/master/strings.pdf){target="_blank"}

[GitHub - tidyverse/stringr](https://github.com/tidyverse/stringr){target="_blank"}

[14 Strings | R for Data Science](https://r4ds.had.co.nz/strings.html){target="_blank"}

[Simple, Consistent Wrappers for Common String Operations • stringr](https://stringr.tidyverse.org/){target="_blank"}
